---
title: "Motus Station Report"
author: "Amie MacDonald and Lucas Berrigan - Motus"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: 
    fig_width: 12
    fig_height: 8
    toc: no
  html_document: 
    fig_width: 12
    fig_height: 8
    toc: no
  mainfont: "Helvetica"
urlcolor: blue
---


<style>


.issue, .no-issue {
  border:solid 1px #000;
  font-weight:bold;
  padding:10px;
  text-align:center;
  text-decoration:no-decoration;
}
.issue {
  opacity:0.8;
}
.issue:hover {
  cursor:pointer;
  opacity:1;
  text-decoration:no-decoration;
}
.warning {
  background-color:#EEE;
  color:#000;
}
.urgent {
  background-color:#FF0;
  color:#A00;
}
.no-issue {
  background-color:#EFE;
}


</style>



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width=20, fig.height=14) 
```

```{r, code = readLines("downloadData.R")}

```

```{r load, include=FALSE}


# Load Libraries
library(tidyverse)
library(rworldmap)
library(rworldxtra)
library(motus)
library(devtools)
library(lubridate)
library(scales)
library(knitr)
library(mapdata)
library(fuzzyjoin)
library(oce)
library(plotly)
library(ggmap)


# Set session time zone to GMT
Sys.setenv(tz = "GMT")

#### Folder where data is stored
#data.dir <- 'C:/Users/Lucas Berrigan/Documents/Data/'
#### 
      

Sys.setenv('motus_userAuth' = '')

message(paste0("Preparing ", ifelse(printable, "", "non-"), "printable version."))

message(paste0("Downloading data from ", data.dir))

downloadMotus_recvData(site.name = receiver.deploy.name, projID = projectID, dir = data.dir)

recv.df <- read.csv(paste0(data.dir, 'receiver-deployments.csv')) %>%
  mutate(tsStart = as.POSIXct(tsStart, origin = '1970-01-01'),
         tsEnd = as.POSIXct(tsEnd, origin = '1970-01-01'))

select.recv.df <- recv.df %>% 
  filter(grepl(receiver.deploy.name, deploymentName, T), recvProjectID == projectID | is.na(projectID))

receiver <- unique(select.recv.df$receiverID)

#### 
#### add 'proj' prefix to filenames?
      filename.prefix <- ''

#### 
#### All data before this date will be filtered out - default = as.Date('2000-01-01')
      earliest.date <- as.Date('2000-01-01')
#### 
#### List of tag deploy IDs to omit - default = c(0)
      tags_to_omit <- c(0, 19156, 28051)
#### 
##################################################################
      
      
      

message(paste0(length(which(file.exists(paste(data.dir,receiver,'.motus', sep = '')))), " of ", length(receiver), " files exist!"))


      
      

```


```{r loadData, results = 'hide'}


sql.table.to.df <- function(recv, tblName, data.dir = '') {
#  message(paste0(data.dir, recv, ".motus"))
  sql <- DBI::dbConnect(RSQLite::SQLite(), paste0(data.dir, recv, ".motus"))
#  message(paste0(DBI::dbListTables(sql), collapse = '\n'))
  sql %>% tbl(tblName) %>% collect() %>% as_tibble() %>% mutate(serno = recv)
#  message("Success")
}

df.deps <- bind_rows(lapply(receiver, function(x){sql.table.to.df(x, 'recvDeps', data.dir)})) %>%
  mutate(tsStart = as.POSIXct(tsStart, origin = '1970-01-01'),
         tsEnd = ifelse(is.na(tsEnd), Sys.time(), tsEnd),
         tsEnd = as.POSIXct(tsEnd, origin = '1970-01-01'))

df.deps$name <- factor(df.deps$name, levels = c(unique(filter(df.deps, grepl(receiver.deploy.name, name))$name), unique(filter(df.deps, !grepl(receiver.deploy.name, name))$name)))

df.batches <- bind_rows(lapply(receiver, function(x){sql.table.to.df(x, 'batches', data.dir)})) %>%
  mutate(tsStart = as.POSIXct(tsStart, origin = '1970-01-01'),
         tsEnd = as.POSIXct(tsEnd, origin = '1970-01-01'))

df.activity <- bind_rows(lapply(receiver, function(x){sql.table.to.df(x, 'pulseCounts', data.dir)})) %>%
  mutate(ts = as.POSIXct(as.Date('1970-01-01')) + hours(hourBin),
         date = as.Date(ts),
         ant = factor(ant)) %>%
  filter(as.Date("2010-01-01") < date) %>%
  fuzzy_left_join(
      filter(df.deps, grepl(receiver.deploy.name, name)), 
      by = c('ts' = 'tsStart', 'ts' = 'tsEnd'), 
      match_fun = list(`>=`, `<=`) 
    ) %>%
  filter(!is.na(deployID),
         serno.x == serno.y) %>%
  rename(serno = serno.x) %>%
  select(-serno.y)


#df.activity %>% saveRDS("df.activity.rds")

#df.activity <- readRDS("df.activity.rds")

df.gps <- bind_rows(lapply(receiver, function(x){sql.table.to.df(x, 'gps', data.dir)})) %>%
  mutate(ts = as.POSIXct(ts, origin = '1970-01-01'),
         date = as.Date(ts)) %>%
  filter(as.Date("2010-01-01") < date) %>%
  fuzzy_left_join(
      select(filter(df.deps, grepl(receiver.deploy.name, name)), -serno), 
      by = c('ts' = 'tsStart', 'ts' = 'tsEnd'), 
      match_fun = list(`>=`, `<=`)
    ) %>%
  filter(!is.na(deployID))

# Load the RDS file and clean it up.
df.raw <- paste0(data.dir, filename.prefix, receiver, ".rds") %>%
  map_dfr(readRDS) %>%
  filter(!tagDeployID %in% tags_to_omit,
         grepl(receiver.deploy.name, recvDeployName)
         ) %>%
  mutate(ts = as.POSIXct(ts, origin = '1970-01-01'),
         lat = ifelse(is.na(gpsLat), recvDeployLat, gpsLat),
         lon = ifelse(is.na(gpsLon), recvDeployLon, gpsLon)) %>%
  distinct()



# Remove unwanted tag deployments and dates
df <- df.raw[!duplicated(df.raw$hitID),] %>% 
  filter(runLen > 2, !is.na(tagProjID), ts > earliest.date) %>%
  mutate(speciesEN = ifelse(is.na(speciesEN), ifelse(is.na(tagDeployTest), "No data", "Test tag"), speciesEN)) %>% 
  filter(tagProjID %in% c(191, 280, 349)) # get rid of false positives at Blackie Spit

tower.lat <- median(df$lat, na.rm = T)
tower.lon <- median(df$lon, na.rm = T)

station.ll <- df.deps %>% 
  mutate(selected = grepl(receiver.deploy.name, name)) %>%
  filter(selected) %>%
  rename(lat = latitude, lon = longitude) %>%
  summarise(lat = median(lat, na.rm = T), lon = median(lat, na.rm = T)) %>%
  as.list()

station.declination  <- magneticField(tower.lon, tower.lat, median(df$ts))$declination


df.ant <- bind_rows(lapply(receiver, function(x){sql.table.to.df(x, 'antDeps', data.dir)})) %>%
  mutate(bearingTrue = round(bearing - station.declination)) %>%
  rename(bearingMagnetic = bearing) %>%
  select(-serno) %>%
  left_join(df.deps, by = 'deployID') %>%
  mutate(selected = grepl(receiver.deploy.name, name),
         depLen = difftime(tsEnd, tsStart, units = 'secs'),
         name = factor(ifelse(selected, receiver.deploy.name, as.character(name))),
         name = factor(name, levels = c(receiver.deploy.name, levels(name)[which(levels(name) != receiver.deploy.name)])))



df.ant.selected <- df.ant %>%
  filter(selected) %>%
  mutate(port = factor(port),
         tsMean = tsStart + (difftime(tsEnd, tsStart, units = 'secs')/2),
         heightMeters = ifelse(is.na(heightMeters), 0, heightMeters),
         plot.lon.lg = longitude + (sin(pi * (-bearingTrue + 90) / 180)*-0.0125),
         plot.lat.lg = latitude + (cos(pi * (-bearingTrue + 90) / 180)*0.0125),
         plot.lon.sm = longitude + (sin(pi * (-bearingTrue + 90) / 180)*-0.0275),
         plot.lat.sm = latitude + (cos(pi * (-bearingTrue + 90) / 180)*0.0275)) 

  
recv.df <- read.csv(paste0(data.dir, 'receiver-deployments.csv')) %>%
  mutate(tsStart = as.POSIXct(tsStart, origin = '1970-01-01'),
         tsEnd = as.POSIXct(tsEnd, origin = '1970-01-01'))

cols <- c("#89C5DA", "#DA5724", "#74D944", "#CE50CA", "#C0717C", "#CBD588", "#5F7FC7", 
          "#673770", "#D3D93E", "#38333E", "#508578", "#D7C1B1", "#689030", "#AD6F3B", "#CD9BCD", 
          "#D14285", "#6DDE88", "#652926", "#7FDCC0", "#C84248", "#8569D5", "#5E738F", "#D1A33D", 
          "#8A7C64", "#599861")


recv.stats.df <- recv.df %>%
  filter(receiverID %in% receiver,
         grepl(receiver.deploy.name, deploymentName)) %>%
  summarise(min.ts = min(tsStart, na.rm = T), max.ts = max(tsEnd)) %>%
  mutate(max.ts = as.POSIXct(ifelse(is.na(max.ts), Sys.time(), max.ts), origin = '1970-01-01'),
         numDays = round(difftime(max.ts, min.ts, units = 'days')),
         numYears = year(max.ts) - year(min.ts))  

```
# `r receiver.name`

<div style="border:solid 1px #888;background:#FDD;padding:5px;">
#### Disclaimer:
These data have not been inspected for false positives. Broad filters have been applied do reduce the amount of erroneous data, but some may still exist in this report. Any suspicious data can be reported to us here: motus@birdscanada.org
</div>

```{r output, fig.width = 5, fig.height = 4, echo=FALSE}


  # Get lat/lon bounding box around these sites
  latLonBounds <- list((tower.lon + c(-1, 1)),
    (tower.lat + c(-1, 1))
  )

if (F) {
  # Make a new high resolution map
  lakes <- map_data('lakes')
  lakes.df <- fortify(lakes)
  lakes <- NULL
  
  worldMap <- getMap(resolution = "high")
  
  # Connect up all the points so the polygons are closed
  worldMap.df <- fortify(worldMap)
  worldMap <- NULL
  
  
  
  # Make a plain map to show station in context with others nearby in the network.
  recv.df %>%
    filter(latitude > latLonBounds[[2]][1] & latitude < latLonBounds[[2]][2],
           longitude > latLonBounds[[1]][1] & longitude < latLonBounds[[1]][2],
           is.na(tsEnd)) %>% 
  ggplot(aes(longitude, latitude))+
    geom_polygon(data = worldMap.df, aes(long, lat,group=group), fill="#AAAAAA", colour="#000000")+
    geom_polygon(data = lakes.df, aes(long, lat,group=group), fill="#d1dbe5", colour="#000000")+
    geom_point(aes(fill = 'red'),
      shape = 21,
      stroke = 2, 
      size = 2, 
      alpha = 1)+
    geom_point(aes(x = tower.lon, y = tower.lat), shape = 8,
      color = 'black',
      stroke = 2,
      size = 4, 
      alpha = 1)+
    geom_point(aes(x = tower.lon, y = tower.lat, fill = 'yellow'), shape = 21,
      color = 'black',
      stroke = 2, 
      size = 3, 
      alpha = 1)+
    coord_fixed(xlim = latLonBounds[[1]], ylim = latLonBounds[[2]])+
    theme(panel.background = element_rect(fill = '#CCDDFF'),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          legend.title = element_blank(),
          axis.title = element_blank()) +
    scale_fill_manual(values = c('red' = 'red', 'yellow' = 'yellow'), guide = 'legend', labels = c('Other stations', receiver.name))+ 
    guides(fill = guide_legend(reverse = TRUE))
}

#  Load a small map from Google Earth
map.small <- get_map(location = c(lon = tower.lon, lat = tower.lat), zoom = 9, maptype = "hybrid")

# Get list of nearby receivers
nearby.recv.df <- recv.df %>%
  filter(latitude > latLonBounds[[2]][1] & latitude < latLonBounds[[2]][2],
         longitude > latLonBounds[[1]][1] & longitude < latLonBounds[[1]][2],
         is.na(tsEnd))

# Plot a satellite map showing antenna directions in context of other stations
  ggmap(map.small, base_layer = ggplot(data = nearby.recv.df, aes(x = longitude, y = latitude)))+
      geom_point(fill = 'red',
                 shape = 21,
                 stroke = 2, 
                 size = 2, 
                 alpha = 1)+
      geom_point(aes(x = tower.lon, y = tower.lat), 
                 shape = 8,
                 color = 'black',
                 stroke = 2,
                 size = 4, 
                 alpha = 1)+
      geom_point(aes(x = tower.lon, y = tower.lat), 
                 fill = 'yellow', 
                 shape = 21,
                 color = 'black',
                 stroke = 2, 
                 size = 3, 
                 alpha = 1)+
    geom_text(data = filter(df.ant.selected, status == 'active'), aes(x = plot.lon.lg, y = plot.lat.lg, angle = - bearingTrue + 90, color = port), 
              label="-->", size = 7)+
    labs(color = 'Antenna')+
  #    coord_fixed(xlim = latLonBounds[[1]], ylim = latLonBounds[[2]])+
      theme(panel.background = element_rect(fill = '#CCDDFF'),
            panel.grid.major = element_blank(),
            panel.grid.minor = element_blank(),
            axis.text = element_blank(),
            axis.ticks = element_blank(),
            axis.title = element_blank()) +
      scale_fill_manual(values = c('red' = 'red', 'yellow' = 'yellow'), guide = 'legend', labels = c('Other stations', receiver.name))

```

 **First deployed:** `r as.Date(min(select.recv.df$tsStart, na.rm = T))`


 **Number of days receiver has been running:** `r recv.stats.df$numDays`


 **Number of years receiver has been running:** `r recv.stats.df$numYears`
 
 
 **Most recent data:** `r as.Date(max(df.activity$ts, na.rm = T))`


 **Number of antennas:** `r length(unique(df.ant[df.ant$selected,]$port[which(df.ant[df.ant$selected,]$deployID == max(df.ant[df.ant$selected,]$deployID))]))`


 **Number of different animals detected:** `r length(unique(df$tagDeployID))`


 **Number of different species detected:** `r length(unique(df$speciesEN))`


 **Total number of detections over all years:** `r length(unique(df$hitID))`



***

```{r device.history}

basic.colours <- c("red", "green", "blue", "purple", "yellow", "turquoise", "orange")

n.deps <- length(unique(df.ant$name))

cust.cols <- c(hcl(seq(from = 0, to = 360 - (360 / n.deps), length.out = n.deps), 100, 75, alpha = c(1, rep(0.1, n.deps - 1))))

dull.colours <- c("red", 
                  alpha("green", 0.25), 
                  alpha("blue", 0.25), 
                  alpha("purple", 0.25), 
                  alpha("yellow", 0.25), 
                  alpha("turquoise", 0.25), 
                  alpha("orange"))

serno.selected.deps <- df.ant %>%
  group_by(serno) %>%
  summarise(selected.start = as.integer(min(tsStart[which(selected)], na.rm = T)))


```


# Detections

```{r}

df %>%
  filter(runLen > 2, !is.na(tagDeployID)) %>%
  mutate(Year = as.character(year(ts))) %>%
  group_by(Year) %>%
  summarise(`Species` = length(unique(speciesEN)), `Individuals` = length(unique(tagDeployID)), `Detections` = n()) %>%
  kable()

```


```{r, fig.width = 7, fig.height = 4.5}

df.bySpp <- df %>%
  filter(runLen > 2, !is.na(speciesEN)) %>%
  group_by(speciesEN) %>%
  summarise(Hits = n(), Individuals = length(unique(tagDeployID)), Days = length(unique(date(ts))))

  
df.bySpp %>%
  mutate(Species = ifelse(speciesEN %in% arrange(top_n(df.bySpp, 24, Individuals), -Individuals)[1:24,]$speciesEN, speciesEN, 'Other')) %>% 
  group_by(Species) %>%
  summarise(Hits = sum(Hits, na.rm = T), Individuals = sum(Individuals, na.rm = T), Days = max(Days, na.rm = T)) %>%
  ggplot(aes(x = "", y = Individuals, fill = fct_reorder(Species, Individuals))) +
  geom_bar(width = 1, stat = "identity", position='dodge')+
  labs(y = "Number of Individual Animals", x = "", fill = 'Species', title = "Number of individual animals per species")+
  scale_fill_manual(values = cols)+
  theme_bw()+
  theme(panel.border = element_blank())+ 
  guides(fill = guide_legend(reverse = TRUE))

df.bySpp %>% 
  rename(Species = speciesEN,
         Detections = Hits) %>%
  select(-Days) %>%
  arrange(-Individuals) %>%
  kable

```

---
  
```{r, fig.width = 7, fig.height = 4.5}

df.byProj <- df %>%
  filter(runLen > 2, !is.na(tagProjID), !is.na(speciesEN)) %>%
  group_by(tagProjName) %>%
  summarise(Hits = n(), Individuals = length(unique(tagDeployID)), Days = length(unique(date(ts))), Species = length(unique(speciesEN)))

df.byProj %>%
  mutate(Project = as.character(ifelse(tagProjName %in% arrange(top_n(df.byProj, 24, Individuals), -Individuals)[1:24,]$tagProjName, tagProjName, 'Other'))) %>%  
  group_by(Project) %>%
  summarise(Hits = sum(Hits, na.rm = T), Individuals = sum(Individuals, na.rm = T), Days = max(Days, na.rm = T)) %>%
  ggplot(aes(x = "", y = Individuals, fill = fct_reorder(Project, Individuals))) +
  geom_bar(width = 1, stat = "identity", position = 'dodge') +
 # coord_polar("y", start = 0)+
  labs(y = "Number of Individual Animals", x = "", title = "Number of individuals animals per project", fill = 'Project ID')+
  scale_fill_manual(values = cols)+
  theme_bw()+
  theme(panel.border = element_blank())+ 
  guides(fill = guide_legend(reverse = TRUE))

df.byProj  %>% 
  rename(`Project name` = tagProjName,
         Detections = Hits) %>%
  select(-Days) %>%
  arrange(-Individuals) %>%
  kable
```






```{r, fig.width = 7, fig.height = 4.5}

top5 <- df %>% 
  filter(runLen > 2, !is.na(tagDeployID)) %>%
  group_by(speciesEN) %>%
  summarise(nInd = length(unique(tagDeployID)), nHits = n()) %>%
  mutate(propHits = nHits/nrow(df)) %>%
  arrange(desc(nHits))

top <- top5 %>%
  mutate(Species = factor(ifelse(nHits %in% top5[1:ifelse(nrow(top5) == 5, 5, 4),]$nHits, speciesEN, 'Other'))) %>%
  select(speciesEN, Species)

levels(top$Species) <- c(levels(top$Species)[!grepl('Other', levels(top$Species))][order(levels(top$Species)[!grepl('Other', levels(top$Species))])], "Other")


plot.binwidth <- as.integer(round(difftime(range(df$ts)[2], range(df$ts)[1])/100))

plot.binwidth <- ifelse(plot.binwidth == 0, 1, plot.binwidth)

df %>%
  filter(runLen > 2, !is.na(tagDeployID)) %>%
  mutate(date = date(ts)) %>%
  group_by(tagDeployID, speciesEN, date) %>%
  summarise(nHits = n()) %>% 
  left_join(top, by = 'speciesEN') %>%
  ggplot(aes(date, fill = Species)) +
  geom_histogram(binwidth = plot.binwidth)+ 
  scale_fill_manual(values = cols)+
  labs(x = 'Date', y = 'Number of Individuals', fill = 'Species', title = 'Number of individuals detected each day')+
  theme_minimal()
```


---


```{r, fig.width = 7, fig.height = 4.5}

df %>%
  filter(runLen > 2, !is.na(tagDeployID)) %>%
  mutate(Hour = hour(as.POSIXct(ts, origin = '1970-01-01'))-8,
         Hour = ifelse(Hour < 0, Hour + 24, Hour)) %>%
  left_join(top, by = 'speciesEN') %>%
  ggplot(aes(Hour, fill = Species)) +
  geom_histogram(bins = 24)+ 
  coord_polar("x", start = 0)+
  labs(x = 'Hour', y = 'Number of detections', fill = 'Species', title = 'Number of detections by time of day')+
  theme_minimal()
 

```


## List of all individuals detected at this station
Click on Tag ID to learn more about the individual or 'view track' to see its animated flight path.


```{r}



allInd.list <- df %>% 
  filter(runLen > 2, !is.na(tagDeployID)) %>%
  group_by(tagDeployID, speciesEN) %>%
  summarise(`Detections` = n(), 
            `Days` = length(unique(date(ts))), 
            `First detected` = min(date(ts)), 
            `Last detected` = max(date(ts))) %>%
  ungroup() 

if (!printable) {
  allInd.list <- allInd.list %>%
    mutate(`   ` = paste("[View Track]", "(https://motus.org/data/track?tagDeploymentId=", tagDeployID, ")", sep = ''),
         tagDeployID = paste("[", tagDeployID, "]", "(https://motus.org/data/tagDeployment?id=", tagDeployID, ")", sep = ''))
}

allInd.list %>%
  arrange(desc(`Last detected`)) %>%
  rename(`Tag ID` = tagDeployID, Species = speciesEN) %>%
  kable()

```
